API design for machine learning software: experiences from the scikit-learn project

### 2013-06-22 paper.tex
minor copyedit

From its core, the library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a numeric array data type and proper
array computing primitives, while SciPy \citep{varoquaux2013scipy} extends it
further by making common numerical operations available, either by implementing
those in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
implementations. In a similar vein, \sklearn extends Python with a set of
composable machine learning operations.

@@

\begin{description}
  \item[Consistency.]
       All objects (basic or composite) share a consistent interface composed of
       a limited set of methods. This interface is documented in a consistent
       manner for all objects.
  \item[Inspection.]
       Constructor parameters and parameter values determined by learning
       algorithms are stored and exposed as public attributes.
  \item[Non-proliferation of classes.]
       Learning algorithms are the only objects to be of custom classes.
       Datasets and hyper-parameter names and values are represented as standard
       Python, NumPy or SciPy classes rather than new ones whenever feasible.
       This keeps \sklearn easy to use and easy to combine with other libraries.
  \item[Composition.]
       Many machine learning tasks are expressible
       as sequences or combinations of transformations to data.
       Some learning algorithms are also naturally viewed
       as meta-algorithms parametrized on other algorithms.
       Whenever possible, those algorithms are implemented and composed from
       existing building blocks.
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       a default and appropriate value is defined by the library.
       The default value should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}

@@

In \sklearn, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While those may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as those used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
on all NumPy and SciPy vector operations while keeping
the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices.
From a practical point of view, these formats also provide a bunch of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

@@

In \sklearn, classical learning algorithms are not the only objects to be
implemented as estimators. For example, preprocessing routines (e.g., scaling of
features) or feature extraction techniques (e.g., vectorization of text
documents) also implement the same interface. Even those estimators that are
entirely stateless and therefore do not require the \texttt{fit} method to
perform any useful work implement the estimator interface. As we will illustrate
in the next sections, this design pattern is indeed of prime importance for
consistency, composition and model selection reasons.

@@

Since meta-estimators require users to construct nested objects,
the decision to implement a meta-estimator
rather than fold the behavior it implements
into existing estimators classes
is always based on a trade-off between genericity and ease of use.
Relating to the example just given,
all \sklearn classifiers are designed to do multiclass classification
and the use of the \texttt{multiclass} module
is only necessary in advanced use cases.

@@

In \sklearn, model selection is supported in two distinct meta-estimators,
\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as inputs an
estimator (basic or composite), whose hyper-parameters must be optimized, and a
hyper-parameter grid, which defines the values to search. Optionally, they can
also take a cross-validation scheme and a score function.  The library provides
various such cross-validation schemes, including K-Fold (default), Leave-One-Out
and Bootstrap. The score function used by default is the \texttt{score} method
provided by the estimator. However, the library provides a large variety of
score functions that the user can choose from,
including accuracy, AUC, $F_1$ score (classification),
$R^2$ score and mean squared error (regression).

@@

To ease code reuse, simplify implementation and skip the introduction of
superfluous classes, the principle of \textit{duck typing} is exploited
throughout the codebase. This technique allows both for extensibility and
flexibility: as long as an estimator follows the API and conventions
outlined in Section~\ref{sec:core-api}, then it can be used in lieu of a
built-in estimator (e.g., it can plugged into pipelines or grid search).
That is, estimators are defined only by interface, not by inheritance,
and external developers are not forced to inherit from any \sklearn class.

@@

Our rule of thumb is that user code should not be tied to \sklearn---which is a
\textit{library}, and not a \textit{framework}. This principle indeed avoids a
well-known problem with object-oriented design, which is that users wanting a
``banana'' should not get ``a gorilla holding the banana and the entire jungle''
(J.~Armstrong, cited by \citealp[p.~213]{seibel2009coders}).  That is, programs
using \sklearn should be easily reusable with other machine learning toolkits or
in other contexts, because they don't need to be intimately tied to it.

>>>>

From its core, the library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a numeric array data type and proper
array computing primitives, while SciPy \citep{varoquaux2013scipy} extends it
further by making common numerical operations available, either by implementing
these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
implementations. In a similar vein, \sklearn extends Python with a set of
composable machine learning operations.

@@

\begin{description}
  \item[Consistency.]
       All objects (basic or composite) share a consistent interface composed of
       a limited set of methods. This interface is documented in a consistent
       manner for all objects.
  \item[Inspection.]
       Constructor parameters and parameter values determined by learning
       algorithms are stored and exposed as public attributes.
  \item[Non-proliferation of classes.]
       Learning algorithms are the only objects to be of custom classes.
       Datasets and hyper-parameter names and values are represented as standard
       Python, NumPy or SciPy classes rather than new ones whenever feasible.
       This keeps \sklearn easy to use and easy to combine with other libraries.
  \item[Composition.]
       Many machine learning tasks are expressible
       as sequences or combinations of transformations to data.
       Some learning algorithms are also naturally viewed
       as meta-algorithms parametrized on other algorithms.
       Whenever possible, such algorithms are implemented and composed from
       existing building blocks.
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       a default and appropriate value is defined by the library.
       The default value should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}

@@

In \sklearn, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While these may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as the ones used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
on all NumPy and SciPy vector operations while keeping
the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices.
From a practical point of view, these formats also provide a bunch of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

@@

In \sklearn, classical learning algorithms are not the only objects to be
implemented as estimators. For example, preprocessing routines (e.g., scaling of
features) or feature extraction techniques (e.g., vectorization of text
documents) also implement the same interface. Even processing steps that are
entirely stateless and therefore do not require the \texttt{fit} method to
perform any useful work implement the estimator interface. As we will illustrate
in the next sections, this design pattern is indeed of prime importance for
consistency, composition and model selection reasons.

@@

Since meta-estimators require users to construct nested objects,
the decision to implement a meta-estimator
rather than fold the behavior it implements
into existing estimators classes
is always based on a trade-off between generality and ease of use.
Relating to the example just given,
all \sklearn classifiers are designed to do multiclass classification
and the use of the \texttt{multiclass} module
is only necessary in advanced use cases.

@@

In \sklearn, model selection is supported in two distinct meta-estimators,
\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as inputs an
estimator (basic or composite), whose hyper-parameters must be optimized, and a
hyper-parameter grid, which defines the values to search. Optionally, they can
also take a cross-validation scheme and a score function.  The library provides
various such cross-validation schemes, including $k$-fold (default),
leave-one-out and bootstrapping.
The score function used by default is the \texttt{score} method
provided by the estimator. However, the library provides a large variety of
score functions that the user can choose from,
including accuracy, AUC, $F_1$ score (classification),
$R^2$ score and mean squared error (regression).

@@

To ease code reuse, simplify implementation and skip the introduction of
superfluous classes, the principle of \textit{duck typing} is exploited
throughout the codebase. This technique allows for both extensibility and
flexibility: as long as an estimator follows the API and conventions
outlined in Section~\ref{sec:core-api}, then it can be used in lieu of a
built-in estimator (e.g., it can be plugged into pipelines or grid search).
That is, estimators are defined only by interface, not by inheritance,
and external developers are not forced to inherit from any \sklearn class.

@@

Our rule of thumb is that user code should not be tied to \sklearn---which is a
\textit{library}, and not a \textit{framework}. This principle indeed avoids a
well-known problem with object-oriented design, which is that users wanting a
``banana'' should not get ``a gorilla holding the banana and the entire jungle''
(J.~Armstrong, cited by \citealp[p.~213]{seibel2009coders}).
That is, programs using \sklearn should not be intimately tied to it,
so that their code can be reused with other toolkits or in other contexts.
